\section{Paper Summary}

The paper presents \textit{NeuroPathX}\cite{wang2024neuropathx}, an explainable deep learning framework designed to model interactions between genetic variation and brain structure in neurological disorders. The central methodological challenge is multimodal imaging-genetics integration under two competing requirements: capturing complex nonlinear associations while maintaining interpretability. This problem is particularly relevant in neurological diseases such as Autism Spectrum Disorder (ASD) and Alzheimer's Disease (AD), which are highly heterogeneous and polygenic, meaning that disease risk arises from the combined effect of many genetic variants interacting with brain structure.

\paragraph{Positioning with respect to the state of the art.}
Existing imaging-genetics approaches generally fall into two categories. Classical methods rely on linear multivariate models linking single-nucleotide polymorphisms (SNPs) to imaging phenotypes, such as canonical correlation analysis (CCA), which limits their ability to capture nonlinear interactions. More recent deep learning approaches can model complex relationships but often lack interpretability. Latent-fusion architectures combine modality-specific embeddings but may obscure direct relationships between genes and brain regions \cite{ghosal2021gmind}. Early-fusion attention-based models attempt to explicitly model cross-modal interactions but do not enforce stability or consistency in the learned associations \cite{jaume2024survpath}. NeuroPathX addresses these limitations by introducing biologically structured genetic inputs and regularizing the attention mechanism to produce interpretable pathway-brain associations.

\paragraph{Main contributions.}
The work makes three main contributions. First, genetic variation is aggregated into biologically meaningful pathway-level representations. Disease-associated SNPs obtained from genome-wide association studies (GWAS) are mapped to genes and combined using reported effect sizes, and gene-level signals are subsequently aggregated into KEGG biological pathways. Second, the model introduces a pathway-guided cross-attention mechanism that explicitly models interactions between pathway-level genetic features and region-of-interest (ROI) imaging measurements. Third, the framework incorporates two regularization terms applied directly to the attention matrix: a sparsity loss encouraging selective pathway-brain interactions and a pathway similarity loss promoting cohort-level consistency in the pathways associated with the disorder.

\paragraph{Methodology.}
Genetic preprocessing follows a polygenic risk score pipeline \cite{unknown}. For each subject, SNP genotypes are encoded as allele counts and weighted using GWAS effect sizes. SNPs located within $\pm 50$ kb of a gene are aggregated to produce gene-level scores, which are then summed across genes belonging to the same KEGG pathway \cite{10.1093/nar/28.1.27}, yielding a pathway vector of size $N_p=177$. Structural MRI features are extracted from predefined brain regions of interest using FreeSurfer \cite{FISCHL2012774}, producing anatomical measurements including volume, surface area, cortical thickness, and thickness variability. The core model employs cross-attention in which pathway embeddings act as queries over ROI features, producing an attention matrix representing pathway-brain interactions. In contrast to standard transformer attention, softmax normalization is replaced by a ReLU-SoftSign activation, allowing pathways to have zero influence on brain regions. The resulting pathway-guided imaging representations are fed into a multilayer perceptron for disease classification. The training objective combines weighted cross-entropy with a KL-based sparsity penalty on attention weights and an $\ell_1$ pathway similarity constraint.

\paragraph{Validation and results.}
The method is evaluated using 10-fold cross-validation on two datasets: the ACE dataset for ASD ($N=165$) and the ADNI dataset for AD ($N=438$). Performance is assessed using accuracy, sensitivity, specificity, and AUC. NeuroPathX achieves the highest accuracy and specificity on both datasets and outperforms ablated variants that remove the proposed regularization terms. Interpretability is assessed by analyzing the learned attention matrices, whose averaged pathway-ROI interactions are reported to correspond to known biological processes implicated in ASD and AD.